---
title: "Full Kidney Analysis"
author: "Omid Sajjadi"
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: yes
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,       # Hides the code
  warning = FALSE,    # Hides warnings
  message = FALSE,    # Hides messages (like package loading)
  results = 'hide'    # Hides printed output
)
```

```{r load new packages}

if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("cowplot")) {install.packages("cowplot"); require("cowplot")}
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
if (!require("plotly")) {install.packages("plotly"); require("plotly")}
if (!require("knitr")) {install.packages("knitr"); require("knitr")}

if (!require("here")) {install.packages("here"); require("here")}

```

## Question 1
### 1-1: What filtering parameters did you choose? Why?

Here the data was filtered via this reasoning: Cells with <500 genes per cell were excluded this is because these cells are likely dead or an empty droplet. Further, cells with more than 4000 genes were also removed this voided doublets.Furthermore, cell containing <1000 RNA molecules were also excluded, this took place to avoid dying cells. Moreover, to avoid doublets even more, cells with more than 10% mitochondrial transcripts were also excluded.

```{r, eval=TRUE, fig.align = 'center'}
Dataset <- Read10X(data.dir = here("/Users/omidsajjadi/Desktop/BTG/Coding/Week 2 - Rmarkdown and the Whole Kidney/Week 2 Materials/Dataset"))
Dataset <- CreateSeuratObject(counts = Dataset, project = "Dataset", min.cells = 3, min.features = 200)
Dataset[["percent.mt"]] <- PercentageFeatureSet(Dataset, pattern = "^mt-")
p1 <- VlnPlot(Dataset, features = c("nFeature_RNA"), pt.size = .4) + theme(axis.title.x = element_blank()) + theme(legend.position = 'none') + theme(axis.text.x = element_text(angle = 0, hjust = .5))
p2 <- VlnPlot(Dataset, features = c("nCount_RNA"), pt.size = .4) + theme(axis.title.x = element_blank())  + theme(legend.position = 'none') + theme(axis.text.x = element_text(angle = 0, hjust = .5))
p3 <- VlnPlot(Dataset, features = c("percent.mt"), pt.size = .4) + theme(axis.title.x = element_blank())  + theme(legend.position = 'none') + theme(axis.text.x = element_text(angle = 0, hjust = .5))
```

```{r, eval=TRUE, fig.align = 'center'}
Dataset_Filtered <- subset(Dataset, subset = nFeature_RNA > 500 & nFeature_RNA < 4000 & nCount_RNA > 1000 & percent.mt < 10)
p11 <- VlnPlot(Dataset_Filtered, features = c("nFeature_RNA"), pt.size = 0) + theme(axis.title.x = element_blank()) + theme(legend.position = 'none') + theme(axis.text.x = element_text(angle = 0, hjust = .5)) 
p12 <- VlnPlot(Dataset_Filtered, features = c("nCount_RNA"), pt.size = 0) + theme(axis.title.x = element_blank())  + theme(legend.position = 'none') + theme(axis.text.x = element_text(angle = 0, hjust = .5))
p13 <- VlnPlot(Dataset_Filtered, features = c("percent.mt"), pt.size = 0) + theme(axis.title.x = element_blank())  + theme(legend.position = 'none') + theme(axis.text.x = element_text(angle = 0, hjust = .5))

cowplot::plot_grid(p1, p2, p3, p11, p22, p33, ncol = 3, nrow = 2, scale = .85) + draw_label("Filtered Data", y = .95, size = 22, fontface = "bold")
```

## 1-2: How many cells do they filter out? 

```{r cell-counts, eval=TRUE, fig.align = 'center'}
before <- ncol(Dataset)
after <- ncol(Dataset_Filtered)
removed <- before - after

txt1 <- paste("**Number of cells before filtering:**", before)
txt2 <- paste("**Number of cells after filtering:**", after)
txt3 <- paste("**Cells removed during filtering:**", removed)
```

`txt1`
`txt2`
`txt3`

## Normalizing, Scaling Data, and PCA

X-axis (Average Expression): How much a gene is expressed overall in all cells. Y-axis (Standardized Variance): How much the gene’s expression changes between cells.

```{r, echo= FALSE, fig.align='center'}
Dataset_Filtered <- NormalizeData(Dataset_Filtered, normalization.method = "LogNormalize", scale.factor = 10000)
Dataset_Filtered <- FindVariableFeatures(Dataset_Filtered, selection.method = "vst", nfeatures = 2000)

top10 <- head(VariableFeatures(Dataset_Filtered), 10)
plot1 <- VariableFeaturePlot(Dataset_Filtered)
plot2 <- LabelPoints(plot = plot1, points = top10, repel = TRUE)
plot2 + theme(legend.position = 'none')
plot2

gene_count <- nrow(Dataset_Filtered)
```
This dataset contains `r gene_count` genes.

Below, in the scree bar and the elbow bar we can see that the elbow is much longer, or in other words the bars decline is smooth, this is an indicator that we have much more cell types. This shows that our dataset is indeed a whole kidney.

```{r, echo=FALSE, fig.align='center'}
all.genes <- rownames(Dataset_Filtered)
Dataset_Filtered <- ScaleData(Dataset_Filtered, features = all.genes)
Dataset_Filtered <- RunPCA(Dataset_Filtered, features = VariableFeatures(object = Dataset_Filtered))

print(Dataset_Filtered[["pca"]], dims = 1:5, nfeatures = 5)
pca_scores <- Embeddings(Dataset_Filtered, reduction = "pca")
```

```{r fig.align='center', eval=FALSE}
ElbowPlot(Dataset_Filtered) + labs(title = "Elbow Plot") + theme_minimal() + theme(axis.text.x = element_text(angle = 45, hjust = 0.1))

stdev <- Dataset_Filtered[["pca"]]@stdev
pct_var <- (stdev^2) / sum(stdev^2) * 100
pc_df <- data.frame(
  PC = factor(paste0("PC", 1:length(pct_var)), 
  levels = paste0("PC", 1:length(pct_var))),
  Variance = pct_var,
  Label = sprintf("%.1f%%", pct_var)
  )
#Plot as a bar plot
ggplot(pc_df[1:25, ], aes(x = PC, y = Variance)) +
  geom_bar(stat = "identity", fill = "steelblue") +
  geom_text(aes(label = Label), vjust = -0.3, size = 2) +  # label above bar
  labs(title = "Scree Plot (Bar)", y = "% Variance Explained", x = "Principal Components") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 0.1))
```

This is a Dimplot using PC1 and PC2 is not a good plot and indicates the need for UMAP.

```{r, fig.align='center'}
DimPlot(Dataset_Filtered, reduction = "pca")
```


## Question 2 - UMAP and Cluster Visualization
### What filtering parameters did you choose? Why?  
To determine the optimal parameters for UMAP visualization, I conducted a series of tests using various combinations of principal component (PC) dimensions and clustering resolution values. Through this process, **I identified PC dimensions 1–30 with a resolution of 2.0 as the most effective.** These parameters produced the clearest separation between cell types, resulting in well-defined and biologically meaningful clusters.

```{r, error=FALSE, fig.align='center', message=FALSE, warning=FALSE, results=FALSE, echo=FALSE}
Dataset_Filtered4 <- FindNeighbors(Dataset_Filtered, dims = 1:30)
Dataset_Filtered4 <- FindClusters(Dataset_Filtered4, resolution = 2)
Dataset_Filtered4 <- RunUMAP(Dataset_Filtered4, dims = 1:30)
p4 <- DimPlot(Dataset_Filtered4, reduction = "umap", label = TRUE)
p4
```
```{r, error=FALSE, fig.align='center', message=FALSE, warning=FALSE, results=FALSE, echo=TRUE}
Dataset_Filtered4[["RNA"]] <- JoinLayers(Dataset_Filtered4[["RNA"]])
Data.markers <- FindAllMarkers(Dataset_Filtered4, only.pos = TRUE, min.pct = 0.5, logfc.threshold = 1) # I can speed up this by increasing threshold and pct
Data.markers %>% group_by(cluster) %>% top_n(n = 5, wt = avg_log2FC)
top5 <- Data.markers %>% group_by(cluster) %>% top_n(5, avg_log2FC)
p44 <- DoHeatmap(Dataset_Filtered4, features = top5$gene) 
p44
```

### How many clusters did you get? 
I was able to get 37 cluster, but there are doublets in there and I would need to identify those doublets and remove them to achieve a higher resolution UMAP.

```{r, fig.width= 20, fig.height= 10, echo=FALSE}
markers.to.plot1 <- c("Lrp2",         # PTS1, PTS2, PTS3
                      "Slc5a12",      # PTS1
                      "Slc13a3",      # PTS2
                      "Slc16a9",      # PTS3
                      "Epha7",        # DTL
                      "Cryab",        # DTL
                      "Slc12a1",      # TAL1, TAL2, MD
                      "Cldn16",       # TAL1
                      "Cldn10",       # TAL2
                      "Nos1",         # MD
                      "Slc12a3",      # DCT1, DCT2
                      "Egf",          # DCT1
                      "Slc8a1",       # CNT, DCT2
                      "Aqp2",         # PC
                      "Slc4a1",       # ICA
                      "Kit",          # ICA
                      "Slc26a4",      # ICB
                      "Nphs1",        # PODO
                      "Ncam1",        # PEC
                      "Upk1b",        # URO
                      "Lef1",         # ECG
                      "Flt1",         # ECP
                      "Emcn",         # ECP
                      "Lyve1",        # LYMPH
                      "Kdr",          # LYMPH
                      "Pdgfra",       # FIB
                      "Pdgfrb",       # FIB, MES, VSMC
                      "Piezo2",       # MES
                      "Acta2",        # VSMC
                      "Ptprc",        # LYMPHO, MACRO
                      "Skap1",        # LYMPHO
                      "Cd74"         # MACRO
)
 
p5 <- DotPlot(Dataset_Filtered4,
features = markers.to.plot1,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip() +
  labs(x = "Gene Features", y = "Clusters")
 
p5
```



```{r, echo=TRUE, eval=FALSE}
p6 <- as.data.frame(p5$data)
 
maybe_doublets <- p6 %>%
  group_by(id) %>%
  filter(all(avg.exp.scaled <= 2.0)) %>%
  distinct(id) %>%
  pull(id)
 
maybe_doublets
 
# Create color palette: grey for all, color for maybe_doublets
all_clusters <- levels(Dataset_Filtered4)
cluster_colors <- setNames(rep("grey80", length(all_clusters)), all_clusters)
cluster_colors[maybe_doublets] <- "red"
 
# Get UMAP coordinates and cluster assignments
umap_data <- Embeddings(Dataset_Filtered4, "umap") %>%
  as.data.frame() %>%
  mutate(cluster = Idents(Dataset_Filtered4))
 
# Compute cluster centroids
cluster_centroids <- umap_data %>%
  group_by(cluster) %>%
  summarise(UMAP_1 = mean(umap_1), UMAP_2 = mean(umap_2))
 
# Keep only centroids for maybe_doublets
label_data <- cluster_centroids %>%
  filter(cluster %in% maybe_doublets)
 
# Plot
DimPlot(Dataset_Filtered4, reduction = "umap", cols = cluster_colors) +
  ggrepel::geom_text_repel(
    data = label_data,
    aes(x = UMAP_1, y = UMAP_2, label = cluster),
    size = 4,
    color = "black",
    fontface = "bold",
    max.overlaps = Inf,
    box.padding = 0.5,
    point.padding = 0.5
  )
```



```{r, echo=TRUE, eval=FALSE}
# Apply the name "SINGLET" or "DOUBLET" to a cluster based on the multidimensional dot plot
 
doublet_clusters <- c(20,21,27)
 
# Annotate metadata
Dataset_Filtered4@meta.data <- Dataset_Filtered4@meta.data %>%
  mutate(subclass.doublet = if_else(
    seurat_clusters %in% doublet_clusters, "DOUBLET", "SINGLET"
  ))
 
# Change the identities to align with new subclass names and make umap
Idents(Dataset_Filtered4) <- Dataset_Filtered4@meta.data$subclass.doublet
 
DimPlot(Dataset_Filtered4, reduction = "umap") +
  ggtitle("Cell Clusters in the Kidney - Singlet vs. Doublet Clustering")
 
# Clean out the doublets manually that were identified
seurat.obj.f_cleaned <- subset(Dataset_Filtered4, idents = "DOUBLET", invert = TRUE)
 
# Double check that the identity was removed
table(Idents(seurat.obj.f_cleaned))
```


```{r, echo=TRUE, eval=FALSE}

#code for reclustering post doublet detection
library(sctransform)
seurat.obj.f_cleaned <- SCTransform(seurat.obj.f_cleaned) %>%
    RunPCA() %>%
    FindNeighbors(dims = 1:30) %>%
    FindClusters(resolution = 2, verbose = FALSE) %>%
    RunUMAP(dims = 1:30)
 
DimPlot(seurat.obj.f_cleaned, reduction = "umap", label = TRUE)
```



```{r, echo=TRUE, eval=FALSE}
p4 <- p4 + theme(legend.position = "none") + ggtitle("Res 0.1, PC 1:10")
p5 <- p5 + theme(legend.position = "none") + ggtitle("Res 0.1, PC 1:15")
p6 <- p6 + theme(legend.position = "none") + ggtitle("Res 0.15, PC 1:15")
p7 <- p7 + theme(legend.position = "none") + ggtitle("Res 0.2, PC 1:15")
plot_grid(p4, p5, p6, p7, ncol = 2, nrow = 2)

p44 <- p44 + theme(legend.position = "none") + ggtitle("Res 0.1, PC 1:10")
p55 <- p55 + theme(legend.position = "none") + ggtitle("Res 0.1, PC 1:15")
p66 <- p66 + theme(legend.position = "none") + ggtitle("Res 0.15, PC 1:15")
p77 <- p77 + theme(legend.position = "none") + ggtitle("Res 0.2, PC 1:15")
plot_grid(p44, p55, p66, p77, ncol = 2, nrow = 2)
```

```{r, echo=TRUE, eval=FALSE}
p77
```

## Understanding CLusters

```{r, echo=TRUE, eval=FALSE}
VlnPlot(Dataset_Filtered2, 
        features = c("Nphs2", "Pecam1", "Slc12a3", "Tagln", "Flt1"), 
        pt.size = 0)

```

```{r, echo=TRUE, eval=FALSE}
#cluster 1
FeaturePlot(Dataset_Filtered4, features = c("Gatm"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Slc5a12"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Slc6a19"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Slc5a12"), label = TRUE )

#cluster0
FeaturePlot(Dataset_Filtered4, features = c("Inmt"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Slco1a1"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Gm13936"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Slc13a3"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Slc7a13"), label = TRUE )

#cluster13
FeaturePlot(Dataset_Filtered4, features = c("Thsd7a"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Nphs2"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Nphs1"), label = TRUE )

#cluster2
FeaturePlot(Dataset_Filtered4, features = c("Pecam1"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Exoc3l2"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Ccm2l"), label = TRUE )

#cluster3
FeaturePlot(Dataset_Filtered4, features = c("Slc12a1"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Ptger3"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Enox1"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Nos1"), label = TRUE )

#cluster4
FeaturePlot(Dataset_Filtered4, features = c("Slc5a12"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Slc7a13"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Ghr"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Slc6a18"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Napsa"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Slc22a19"), label = TRUE )

#cluster5
FeaturePlot(Dataset_Filtered4, features = c("Slc12a3"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Trpm6"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Pvalb"), label = TRUE )

#cluster9
FeaturePlot(Dataset_Filtered4, features = c("Slc8a1"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Rhcg"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Slc2a9"), label = TRUE )

#cluster6
FeaturePlot(Dataset_Filtered4, features = c("Fxyd4"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Elf5"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Aqp4"), label = TRUE )

#cluster11
FeaturePlot(Dataset_Filtered4, features = c("Slc26a4"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Hmx2"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Atp6v0d2"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Gm12121"), label = TRUE )

#cluster14
FeaturePlot(Dataset_Filtered4, features = c("Mgl2"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Ccr2"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Ptprc"), label = TRUE )

#cluster7
FeaturePlot(Dataset_Filtered4, features = c("Lama2"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Fhl2"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Pdgfrb"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("C7"), label = TRUE )

#cluster12
FeaturePlot(Dataset_Filtered4, features = c("Gimap5"), label = TRUE )
FeaturePlot(Dataset_Filtered4, features = c("Gimap6"), label = TRUE )

```


## Changing the cluster names and Dotplot

```{r, echo=TRUE, eval=FALSE}
Dataset_Filtered4_copy2 <- UpdateSeuratObject(Dataset_Filtered4)
```

```{r, echo=TRUE, eval=FALSE}

# Redo on full object
Idents(Dataset_Filtered4_copy2) <- factor(Idents(Dataset_Filtered4_copy2), levels = 0:14)
levels(Dataset_Filtered4_copy2)

new.cluster.ids <- c(
  "PTS3", "PTS1-S2", "Endo", "LOH/MD", "PTS2-S3", "DCT", "CD-PC",
  "Mesangial", "Unknown_8", "CNT", "Unknown_10", "CD-IC/CD-Trans",
  "T cells", "Podo", "Macrophages"
)
levels(Dataset_Filtered4_copy2)
levels(Idents(Dataset_Filtered4_copy2))

names(new.cluster.ids) <- levels(Dataset_Filtered4_copy2)
levels(Dataset_Filtered4_copy2)
levels(Idents(Dataset_Filtered4_copy2))

Dataset_Filtered4_copy2 <- RenameIdents(Dataset_Filtered4_copy2, new.cluster.ids)
levels(Dataset_Filtered4_copy2)
levels(Idents(Dataset_Filtered4_copy2))
# Plot updated UMAP with new cluster names
DimPlot(
  Dataset_Filtered4_copy2, 
  reduction = "umap", 
  label = TRUE, 
  pt.size = 0.5, 
  label.size = 7, 
  repel = FALSE
) + NoLegend()
levels(Dataset_Filtered4_copy2)
levels(Idents(Dataset_Filtered4_copy2))

marker_list <- list(
  "PTS3" = c("Inmt", "Slco1a1", "Gm13936", "Slc13a3", "Slc7a13"),
  "PTS1-S2" = c("Gatm", "Slc5a12", "Slc6a19"),
  "Endo" = c("Pecam1", "Exoc3l2", "Ccm2l"),
  "LOH/MD" = c("Slc12a1", "Ptger3", "Enox1", "Nos1"),
  "PTS2-S3" = c("Slc5a12", "Slc7a13", "Ghr", "Slc6a18", "Napsa", "Slc22a19"),
  "DCT" = c("Slc12a3", "Trpm6", "Pvalb"),
  "CD-PC" = c("Fxyd4", "Elf5", "Aqp4"),
  "Mesangial" = c("Lama2", "Fhl2", "Pdgfrb", "C7"),
  "CNT" = c("Slc8a1", "Rhcg", "Slc2a9"),
  "CD-IC/CD-Trans" = c("Slc26a4", "Hmx2", "Atp6v0d2", "Gm12121"),
  "T cells" = c("Gimap5", "Gimap6"),
  "Podo" = c("Thsd7a", "Nphs2", "Nphs1"),
  "Macrophages" = c("Mgl2", "Ccr2", "Ptprc")
)
levels(Dataset_Filtered4_copy2)
levels(Idents(Dataset_Filtered4_copy2))

Idents(Dataset_Filtered4_copy2) <- "seurat_clusters"


DotPlot(Dataset_Filtered4_copy2, features = marker_list) +
  scale_x_discrete(limits = marker_list) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  ggtitle("DotPlot of Selected Markers by Cell Type") +
  coord_flip()


```






















































```{r, echo=TRUE, eval=FALSE}
if (!require("SingleR")) {
  if (!require("BiocManager")) install.packages("BiocManager")
  BiocManager::install("SingleR")
}
library(SingleR)

if (!require("SingleCellExperiment")) {
  BiocManager::install("SingleCellExperiment")
}
library(SingleCellExperiment)
```

```{r, echo=TRUE, eval=FALSE}
# Load the expression matrix
reference_data <- read.table('/Users/omidsajjadi/Desktop/BTG/Coding/Week 2 - Rmarkdown and the Whole Kidney/Week 2 Materials/Dataset/GSE107585_Mouse_kidney_single_cell_datamatrix.txt', header = TRUE, row.names = 1)

# Confirm the format
head(reference_data)
```

```{r, echo=TRUE, eval=FALSE}
query <- as.SingleCellExperiment(Dataset_Filtered2)
```


```{r, echo=TRUE, eval=FALSE}
# Separate the cluster labels (if needed later)
cluster_labels <- reference_data["Cluster_Number", ]  # optional

# Remove Cluster_Number from expression matrix
expr_matrix <- reference_data[!rownames(reference_data) %in% "Cluster_Number", ]

library(SingleCellExperiment)

# Make sure the expression matrix is numeric
expr_matrix <- as.matrix(expr_matrix)
storage.mode(expr_matrix) <- "integer"

# Create reference SCE object
ref_sce <- SingleCellExperiment(list(counts = expr_matrix))

library(SingleR)

# Example with your dataset as reference (ref_sce) and another Seurat object converted to SCE
# query <- as.SingleCellExperiment(Dataset_Filtered4)
# Option 1: Remove lowly expressed genes
keep_genes <- rowSums(counts(ref_sce) > 1) >= 10
ref_sce <- ref_sce[keep_genes, ]
# Run SingleR
pred <- SingleR(test = query, ref = ref_sce, labels = cluster_labels)  # only if you assign proper labels
```
```{r, echo=TRUE, eval=FALSE}
# LOAD LIBRARIES
if (!require("SingleR")) {
  if (!require("BiocManager")) install.packages("BiocManager")
  BiocManager::install("SingleR")
}
library(SingleR)
library(SingleCellExperiment)

# -- ASSUMPTIONS --
# `reference_data` is a data.frame or matrix with:
#   - rows = genes + 1 row named "Cluster_Number"
#   - columns = reference cells
# `query` is a SingleCellExperiment object (e.g., your Seurat query converted)

# 1. Define cluster to cell type mapping
cluster_map <- c(
  "1" = "Endo",
  "2" = "Podo",
  "3" = "PT",
  "4" = "LOH",
  "5" = "DCT",
  "6" = "CD-PC",
  "7" = "CD-IC",
  "8" = "CD-Trans",
  "9" = "Novel1",
  "10" = "Fib",
  "11" = "Macro",
  "12" = "Neutro",
  "13" = "B lymph",
  "14" = "T lymph",
  "15" = "NK",
  "16" = "Novel2"
)

# 2. Extract cluster numbers (column-wise)
cluster_ids <- as.character(unlist(reference_data["Cluster_Number", ]))

# 3. Drop 'Cluster_Number' row to isolate expression matrix
ref_expr <- reference_data[rownames(reference_data) != "Cluster_Number", ]

# 4. Convert to numeric matrix (may be character if from .txt)
ref_expr <- apply(ref_expr, 2, as.numeric)
rownames(ref_expr) <- rownames(reference_data)[rownames(reference_data) != "Cluster_Number"]
colnames(ref_expr) <- colnames(reference_data)

# 5. Optional filtering to reduce memory usage
keep_genes <- rowSums(ref_expr > 0) > 10
ref_expr <- ref_expr[keep_genes, ]

# 6. Assign labels to reference cells
ref_labels <- cluster_map[cluster_ids]

# 7. Create SingleCellExperiment object for reference
ref_sce <- SingleCellExperiment(
  assays = list(logcounts = ref_expr),
  colData = DataFrame(label = ref_labels)
)

# ------------------------------
# PREP QUERY (ALREADY LOADED)
# ------------------------------

# 8. Extract log-normalized matrix from query
expr_mat <- assay(query, "logcounts")

# 9. Extract query cluster labels (make sure this column exists)
clusters <- colData(query)$seurat_clusters

# 10. Manually aggregate expression per cluster
cluster_means <- function(expr_mat, clusters) {
  unique_clusters <- unique(clusters)
  avg_expr <- sapply(unique_clusters, function(cl) {
    rowMeans(expr_mat[, clusters == cl, drop = FALSE])
  })
  rownames(avg_expr) <- rownames(expr_mat)
  colnames(avg_expr) <- unique_clusters
  avg_expr
}

aggregated_query <- cluster_means(expr_mat, clusters)

# 11. Run SingleR
pred <- SingleR(test = aggregated_query, ref = ref_sce, labels = ref_sce$label)

# 12. Results
head(pred$labels)
table(pred$labels)
```













