---
title: "snRNAseq Analysis Homework 3"
author: "Roman Thomas"
date: "2025-07-09"
output:
  html_document: 
    toc: yes
    toc_depth: 3
    toc_float: true
    number_sections: no
    theme: bootstrap
    df_print: paged
    code_folding: hide
    highlight: pygments
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# Load Packages 
```{r load new packages, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
if (!require("dplyr")) {install.packages("dplyr"); require("dplyr")}
if (!require("Seurat")) {install.packages("Seurat"); require("Seurat")}
if (!require("patchwork")) {install.packages("patchwork"); require("patchwork")}
if (!require("cowplot")) {install.packages("cowplot"); require("cowplot")}
if (!require("ggpubr")) {install.packages("ggpubr"); require("ggpubr")}
if (!require("plotly")) {install.packages("plotly"); require("plotly")}
if (!require("knitr")) {install.packages("knitr"); require("knitr")}
if (!require("htmlwidgets")) {install.packages("htmlwidgets"); require("htmlwidgets")}
if (!require("here")) {install.packages("here"); require("here")}
if (!require("EnhancedVolcano")) {BiocManager::install('EnhancedVolcano'); require("EnhancedVolcano")} # volcano plot
if (!require("tidyverse")) {install.packages("tidyverse"); require("tidyverse")} # for titying up data
if (!require("RColorBrewer")) {install.packages("RColorBrewer"); require("RColorBrewer")} # for color brewer
if (!require("sctransform")) {install.packages("sctransform"); require("sctransform")} # for data normalization
if (!require("glmGamPoi")) {BiocManager::install('glmGamPoi'); require("glmGamPoi")} # for data normalization, sctransform
if (!require("openxlsx")) {install.packages("openxlsx"); require("openxlsx")} # to save .xlsx files
if (!require("SoupX")) {install.packages("SoupX"); require("SoupX")}
if (!require("DoubletFinder")) {BiocManager::install('DoubletFinder'); require("DoubletFinder")}
# remotes::install_github('chris-mcginnis-ucsf/DoubletFinder', force = TRUE)
library(DoubletFinder)
set.seed((12345))
here()
```

# Upload Previous Data 
Upload the RDS file from the previous homework to analyze the doublets. 
```{r load RDS file, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
readRDS("filtered_shotgun_kidney_rptHW2.rds")
```


# 2a. Upload New Data and Remove the Ambient RNA using SoupX
```{r load .h5 files and remove ambient RNA, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
tod = Read10X_h5(here("Dataset", "1_raw_feature_bc_matrix.h5"))
toc = Read10X_h5(here("Dataset", "1_filtered_feature_bc_matrix.h5")) 
sc = SoupChannel(tod,toc)

#Make the Seurat object from the filtered control data
SO <- Read10X_h5(here("Dataset", "1_filtered_feature_bc_matrix.h5")) 
SO <- CreateSeuratObject(counts = SO, project = "Peri-INTACT") #refers to intact skin or tissue not damaged

#Cluster the cells with Seurat
SO <- SCTransform(SO, verbose = F)
SO <- RunPCA(SO, verbose = F)
SO <- RunUMAP(SO, dims = 1:30, verbose = F)
SO <- FindNeighbors(SO, dims = 1:30, verbose = F)
SO <- FindClusters(SO, verbose = T)


meta <- SO@meta.data
umap <- SO@reductions$umap@cell.embeddings
clusters <- setNames(meta$seurat_clusters, rownames(meta))

#Ensure the two following numbers are equal 
length(clusters)
nrow(sc$metaData)

sc <- setClusters(sc, clusters)
sc <- setDR(sc, umap)

#Estimate rho
sc = autoEstCont(sc)

#Clean the data
SO_out = adjustCounts(sc)
 
#Create a new Seurat Object out of the cleaned data
seurat.obj <- CreateSeuratObject(SO_out)
```


## Remove Doublets with the Doublet Detector 
```{r remove doublets, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
VlnPlot(seurat.obj, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)


# Minimal filtering of low quality cells 

seurat.obj.f <- subset(seurat.obj, nFeature_RNA > 500)
VlnPlot(seurat.obj.f, features = c("nFeature_RNA", "nCount_RNA"), ncol = 2)


seurat.obj.f


# Pre-process standard workflow
seurat.obj.f <- NormalizeData(object = seurat.obj.f)
seurat.obj.f <- FindVariableFeatures(object = seurat.obj.f)
seurat.obj.f <- ScaleData(object = seurat.obj.f)
seurat.obj.f <- RunPCA(object = seurat.obj.f)
ElbowPlot(seurat.obj.f, ndims = 40)


# PCs around 15
seurat.obj.f <- FindNeighbors(object = seurat.obj.f, dims = 1:30)
seurat.obj.f <- FindClusters(object = seurat.obj.f, resolution = 1)


seurat.obj.f <- RunUMAP(object = seurat.obj.f, dims = 1:30)
DimPlot(seurat.obj.f, reduction = "umap", label = TRUE)
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.height=10, fig.width=12}
# Creating a dot plot to cross reference the features with the cell clusters present, seeing the average expression and percent expressed 
markers.to.plot1 <- c("Lrp2",         # PT
                      "Slc5a12",      # PT-S1
                      "Slc13a3",      # PT-S2
                      "Slc16a9",      # PT-S3
                      "Havcr1",       # Injured PT
                      "Epha7",        # dTL
                      "Cryab",        # dTL
                      "Cdh13",        # dTL1
                      "Slc14a2",      # dTL2
                      "Slc12a1",      # TAL
                      "Umod",         # TAL, DCT1
                      "Egf",          # TAL, DCT1,
                      "Cldn10",       # TAL
                      "Cldn16",       # TAL
                      "Nos1",         # MD
                      "Slc12a3",      # DCT
                      "Pvalb",        # DCT1
                      "Slc8a1",       # DCT2, CNT
                      "Aqp2",         # PC
                      "Slc4a1",       # IC-A
                      "Slc26a4",      # IC-B
                      "Nphs1",        # Podo
                      "Ncam1",        # PEC
                      "Flt1",         # Endo
                      "Emcn",         # Glom Endo
                      "Kdr",          # Capillary Endo
                      "Pdgfrb",       # Perivascular
                      "Pdgfra",       # Fib
                      "Piezo2",       # Mesangial
                      "Acta2",        # Mural
                      "Ptprc",        # Immune
                      "Cd74",         # Macrophage
                      "Skap1",        # B/T Cells 
                      "Upk1b",        # Uro
                      "Top2a"         # Proliferation
)

DotPlot(seurat.obj.f,
features = markers.to.plot1,
dot.scale = 8,
dot.min = 0,
scale.max = 100,
scale.min = 0,
col.min = -2.5,
col.max = 2.5) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  coord_flip()
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.show='hide', results='hide'}
# Calculate each combination of pN and pK
# I set the PCs to 1:20 just to be safe even though I said 15 earlier 
sweep.res.list_seurat.obj.f <- paramSweep(seurat.obj.f, PCs = 1:20, sct = FALSE) 
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
#Summarize each combination of pN and pK
sweep.stats_seurat.obj.f <- summarizeSweep(sweep.res.list_seurat.obj.f, GT = FALSE) 

#Select the pK that corresponds to max bcmvn to optimize doublet detection
bcmvn_seurat.obj.f <- find.pK(sweep.stats_seurat.obj.f)
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
pK <- bcmvn_seurat.obj.f %>% 
  filter(BCmetric == max(BCmetric)) %>%
  select(pK) 

#See pK in the Values Environment
pK <- as.numeric(as.character(pK[[1]]))
```


```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
# Homotypic Doublet Proportion Estimate -------------------------------------------------------------------------------------
annotations <- seurat.obj.f@meta.data$seurat_clusters  
 
homotypic.prop <- modelHomotypic(annotations)           
homotypic.prop

# 10X Multiplet Rate Table (the doublet ratio is # of cells recovered divided by 125000) https://kb.10xgenomics.com/hc/en-us/articles/360001378811-What-is-the-maximum-number-of-cells-that-can-be-profiled-
 
nrow(seurat.obj.f@meta.data)

nExp_poi <- round(nrow(seurat.obj.f@meta.data) # To calculate cell number
                  /125000              # To calculate the doublet ratio
                  *nrow(seurat.obj.f@meta.data))
nExp_poi

nExp_poi_adj <- round(nExp_poi*(1-homotypic.prop))
```


```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
seurat.obj.f_doublets <- doubletFinder(seurat.obj.f,
                        PCs = 1:20,
                        pN = 0.25,
                        pK = pK,
                        nExp = nExp_poi_adj,
                        reuse.pANN = NULL, sct = FALSE)
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
colnames(seurat.obj.f_doublets@meta.data)[6] <- "pANN"
colnames(seurat.obj.f_doublets@meta.data)[7] <- "DF.class"
head(seurat.obj.f_doublets@meta.data)

table(seurat.obj.f_doublets@meta.data$DF.class)
Week3_doubletcount <-sum(seurat.obj.f_doublets@meta.data$DF.class == "Doublet")

DimPlot(seurat.obj.f_doublets, group.by = "DF.class")

VlnPlot(seurat.obj.f_doublets, "nFeature_RNA", group.by = "DF.class")

VlnPlot(seurat.obj.f_doublets, "nCount_RNA", group.by = "DF.class")
```


## Save an Object with Only the Singlets 
```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
seurat.obj.f_singlets <- subset(seurat.obj.f_doublets, DF.class == "Singlet")
seurat.obj.f_singlets

DimPlot(seurat.obj.f_singlets, reduction = "umap")
```


```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
seurat.obj.f_singlets <- seurat.obj.f_singlets[!grepl("^mt-", rownames(seurat.obj.f_singlets)), ]

#Mito Sanity Check
counts <- GetAssayData(seurat.obj.f_singlets, assay = "RNA")
mito.genes <- grep(pattern = "^mt-", x = rownames(x = counts), value = TRUE) 
mito.genes #should be zero

DimPlot(seurat.obj.f_singlets, reduction = "umap", label = T)
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.height=16, fig.width=22}
seurat.obj.f_singlets.markers <- FindAllMarkers(seurat.obj.f_singlets, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25)

seurat.obj.f_singlets.markers %>%
    group_by(cluster) %>%
    top_n(n = 3, wt = avg_log2FC) -> top3

DoHeatmap(seurat.obj.f_singlets, features = top3$gene) + NoLegend()
```


```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.width=10, fig.height=15}
seurat.obj.f_singlets.markers %>%
    group_by(cluster) %>%
    top_n(n = 2, wt = avg_log2FC) -> top2

VlnPlot(seurat.obj.f_singlets, 
        features = unique(top2$gene),
        stack = TRUE, 
        flip = TRUE,
        pt.size = 0)+
        NoLegend()
```


# 2b. Compare Data from Previous Week to Current Data 
```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
seurat_list <- readRDS("filtered_shotgun_kidney_rptHW2.rds")

seurat.list.doublets <- doubletFinder(seurat_list,
                        PCs = 1:10,
                        pN = 0.25,
                        pK = pK,
                        nExp = nExp_poi_adj,
                        reuse.pANN = NULL, sct = FALSE)
```


```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
colnames(seurat.list.doublets@meta.data)[9] <- "pANN"
colnames(seurat.list.doublets@meta.data)[10] <- "DF.class"
head(seurat.list.doublets@meta.data)

table(seurat.list.doublets@meta.data$DF.class)

DimPlot(seurat.list.doublets, group.by = "DF.class")

VlnPlot(seurat.list.doublets, "nFeature_RNA", group.by = "DF.class")

VlnPlot(seurat.list.doublets, "nCount_RNA", group.by = "DF.class")
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.width==15}
seurat.list.singlets <- subset(seurat.list.doublets, DF.class == "Singlet")
seurat.list.singlets

DimPlot(seurat.list.singlets, reduction = "umap", label=T)

head(seurat_list@meta.data)
Week2_doubletcount <-sum(seurat_list@meta.data$subclass.rpt == "Doublet")
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
doublet_counts <- data.frame(Phase = c("Week 2 Data", "Week 3 Data", "Percentage" ), Cell_Number = c(Week2_doubletcount, Week3_doubletcount, Week2_doubletcount/Week3_doubletcount))

print(doublet_counts)
```

<span style="color:#956f13">I only identified 15% of the doublets last week when looking at the current data.  </span>  


# 2c. Localization of Doublets 
<span style="color:#956f13"> The cells for singlets and doublets usually localize to the same area as where singlets go by themselves. The singlets dominate a majority of the graphs that contain both singlets and doublets, with some doublets being spread across the graph with both. They were spread practically in each cluster.  </span>  

# 3. Supercluster & Identify Cell Types 
```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.height=8}
#Apply names to the clusters based on a more specific level of specificity 
seurat.obj.f@meta.data <- seurat.obj.f@meta.data %>% mutate(subgroup.rpt = dplyr::case_when(
seurat_clusters == 0 ~ "PTS2",
seurat_clusters == 1 ~ "Endo",
seurat_clusters == 2 ~ "PTS1",
seurat_clusters == 3 ~ "PTS2",
seurat_clusters == 4 ~ "PTS2",
seurat_clusters == 5 ~ "PTS3",
seurat_clusters == 6 ~ "PTS1",
seurat_clusters == 7 ~ "DCT",
seurat_clusters == 8 ~ "TAL",
seurat_clusters == 9 ~ "TAL",
seurat_clusters == 10 ~ "DCT2",
seurat_clusters == 11 ~ "PC",
seurat_clusters == 12 ~ "Fib",
seurat_clusters == 13 ~ "dTL",
seurat_clusters == 14 ~ "Glom Endo",
seurat_clusters == 15 ~ "Endo",
seurat_clusters == 16 ~ "dTL2",
seurat_clusters == 17 ~ "PT",
seurat_clusters == 18 ~ "CNT",
seurat_clusters == 19 ~ "TAL",
seurat_clusters == 20 ~ "IC-A",
seurat_clusters == 21 ~ "dTL2",
seurat_clusters == 22 ~ "IC-B",
seurat_clusters == 23 ~ "TAL",
seurat_clusters == 24 ~ "Perivascular",
seurat_clusters == 25 ~ "Uro",
seurat_clusters == 26 ~ "Endo",
seurat_clusters == 27 ~ "Immune",
seurat_clusters == 28 ~ "Podo",
seurat_clusters == 29 ~ "Fib",
seurat_clusters == 30 ~ "Fib",
seurat_clusters == 31 ~ "PT",
seurat_clusters == 32 ~ "dTL",
seurat_clusters == 33 ~ "Endo",
))

#Rename the graph now with the cell clusters identified 
Idents(seurat.obj.f) <- seurat.obj.f@meta.data$subgroup.rpt

DimPlot(seurat.obj.f, reduction = "umap", group.by = "subgroup.rpt", label = TRUE, label.size = 4) +
  ggtitle("Named Cell Clusters in the Kidney - Subgroup Clustering")
```

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.height=7}
#Change the specificty of the cells to a lower level and make another UMAP
seurat.obj.f@meta.data <- seurat.obj.f@meta.data %>%
  mutate(group.rpt = dplyr::case_when(
    subgroup.rpt %in% c("PTS1", "PTS2", "PTS3", "PT") ~ "Proximal Tubule",
    subgroup.rpt %in% c("TAL", "dTL", "dTL2") ~ "Loop of Henle ",
    subgroup.rpt %in% c("DCT", "DCT2", "CNT") ~ "Distal Tubule",
    subgroup.rpt %in% c("PC", "IC-A", "IC-B") ~ "Collecting Duct",
    subgroup.rpt %in% c("Endo", "Glom Endo", "Perivascular") ~ "Vascular",
    subgroup.rpt == "Podo" ~ "Podocytes",
    subgroup.rpt %in% c("Immune", "Fib", "Uro") ~ "Immune System"
  ))

DimPlot(seurat.obj.f, group.by = "group.rpt", label = TRUE, label.size = 4) +
  ggtitle("Named Cell Clusters in the Kidney - Group Clustering") 

saveRDS(seurat.obj.f_singlets, here("Singlets.rds"))
```

# 4. Mesangial Cell Modular Score 
```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE, fig.width=10, fig.height=8}
Mesangial_marker_gene_list <- list(c("Serpine2", "Fhl2", "Des", "Prkca", "Art3", "Nt5e", "Ptgfrb")) # 7 genes from Mesangial Cell Section of Annual Review of Physiology Paper 

seurat.obj.f_singlets <- AddModuleScore(object = seurat.obj.f_singlets, features = Mesangial_marker_gene_list, name = "Mesangial_score")

#Check for score in the metadata

head(seurat.obj.f_singlets@meta.data)

VlnPlot(seurat.obj.f_singlets, "Mesangial_score1")
```
## 4a. 
<span style="color:#956f13">It appears that each cluster actually contains mesangial cells based on the genes specific to mesangial cells.  </span>

## 4b.  
<span style="color:#956f13">There are more cell types than just mesangial cells present in these clusters, as they have all been identified under another cell type, and none of them were identified as mesangial cells.  </span>  

# 5. Forloop for Vln Plot for Top 3 Genes in Each Cluster
```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
clusters <- unique(top3$cluster)

for (clust in clusters) {

gene_list <- top3 %>% filter(cluster == clust) %>% pull(gene)

for (gene in gene_list) {
A <- VlnPlot(seurat.obj.f_singlets, features = gene, group.by = "seurat_clusters") +
         ggtitle(paste("Cluster", clust, "-", gene))
    print(A)
}}
```

# 6. DEG Analysis on Clusters 
```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}
clusters <- unlist(levels(seurat.obj.f_singlets$seurat_clusters))
names(clusters) <- as.numeric(clusters)

DEG.env <- new.env()

ls(DEG.env)
```



# Session Info

```{r, echo=TRUE, warning=FALSE, error=FALSE, message=FALSE}

sessionInfo()

```

